---
title: "Network Analysis of Emotions"
author: "182STG08 박정민"
date: "2019년 3월 28일"
output:
  html_document :
    df_print : paged
---

[Network Analysis of Emotions](https://www.r-bloggers.com/network-analysis-of-emotions/)

* create a visual network of emotions
* collected from 11,000 participants on the emotions they felt at a given moment
* different emotions are highly interconnected, such that one emotion morphs into another and so on  


What is a network?

* a complex set of interrelations between variables
    + `nodes` : variables (in this case, emotions)
    + `edges` : the relationships between the variables  


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
# download data
emotion_raw <- read_csv("https://osf.io/e7uab/download") %>%
  rename(Offense = Ofense,
         Embarrassment = Embarassment)
emotion_raw

```


### Edge data set
```{r, message=FALSE, warning=FALSE}
# id, Hour, Day별로 나타난 감정 분류
emotion_long <- emotion_raw %>%
  gather(emotion_type, value, Pride:Anger) %>%
  arrange(id, Day) %>%
  filter(value == 1) %>%
  select(-value)
emotion_long

emotion_edges <- emotion_long %>%
  mutate(second_emotion = lead(emotion_type)) %>%  # lead : find next value
  rename(first_emotion = emotion_type) %>%
  select(id, Day, Hours, first_emotion, second_emotion) %>%
  group_by(id) %>%
  slice(-length(id))  # id가 바뀔때의 행 제거
emotion_edges

```

### Node data set
```{r, message=FALSE, warning=FALSE}
emotion_nodes <- emotion_long %>%
  count(emotion_type) %>%
  rowid_to_column("id") %>%  # adds a column at the start of the dataframe
  rename(label = emotion_type) %>%
  mutate(valence = ifelse(label %in% c("Awe", "Amusement", "Joy", "Alertness",
                                       "Hope", "Love", "Gratitude", "Pride",
                                       "Satisfaction"),
                          "positive", "negative"))
emotion_nodes

```


### Network data set
```{r, message=FALSE, warning=FALSE}
emotion_network <- emotion_edges %>%
  group_by(first_emotion, second_emotion) %>%
  summarize(weight = n()) %>%
  ungroup() %>%
  select(first_emotion, second_emotion, weight)
emotion_network

edges <- emotion_network %>%
  left_join(emotion_nodes, by = c("first_emotion" = "label")) %>%
  rename(from = id)
edges <- tibble::as.tibble(edges)
tibble:::print.tbl(head(edges))

edges <- edges %>%
  left_join(emotion_nodes, by = c("second_emotion" = "label")) %>%
  rename(to = id) %>%
  select(from, to, weight) %>%
  mutate(weight = ifelse(weight > 4500, 4500, weight))
edges <- tibble::as.tibble(edges)
tibble:::print.tbl(head(edges))


```


### Visualization
```{r, message=FALSE, warning=FALSE, fig.align='center'}
library(tidygraph)
library(ggraph)

network <- tbl_graph(emotion_nodes, edges, directed = TRUE)
network

set.seed(190318)

ggraph(network, layout = "graphopt") +
  geom_edge_link(aes(width = weight, color = scale(weight), alpha = weight), check_overlap = TRUE) +
  scale_edge_color_gradient2(low = "darkgrey", mid = "#00BFFF", midpoint = 1.5, high = "dodgerblue2") +
  scale_edge_width(range = c(.2, 1.75)) +
  geom_node_label(aes(label = label, fill = valence), size = 4) +
  scale_fill_manual(values = c("#FF6A6A", "#43CD80")) +
  theme_graph() +
  theme(legend.position = "none", plot.background = element_rect(fill = "black"))

```


### Using `networkD3`

> VERY IMPORTANT to transform the values to base 0

```{r, message=FALSE, warning=FALSE}
library(networkD3)

nodes_d3 <- emotion_nodes %>%
  mutate(id = id - 1,
         n = (scale(n) + 3)^3)

edges_d3 <- edges %>%
  mutate(from = from - 1, to = to - 1,
         weight = ifelse(weight < 600, 0, log(weight)))

nodes_d3
edges_d3

```


```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=8, fig.align='center'}
forceNetwork(Links = edges_d3, 
             Nodes = nodes_d3, 
             Source = "from", Nodesize = "n",
             Target = "to", NodeID = "label", 
             Group = "valence", Value = "weight", 
             fontFamily = "sans-serif",
             colourScale = JS('d3.scaleOrdinal().domain(["negative", "positive"]).range(["#FF6A6A", "#43CD80"])'),
             opacity = 1, fontSize = 24, linkDistance = 300, linkColour = c("#8DB6CD"),
             arrows = TRUE, zoom = TRUE, bounded = TRUE, legend = TRUE)

```


```{r, message=FALSE, warning=FALSE}
library(networkD3)
titanic <- DALEX::titanic
titanic <- titanic[complete.cases(titanic),]
dim(titanic)
titanic <- titanic %>% 
  mutate(Age=as.factor(ifelse(age<20,'Child','Adult')))

myedge <- titanic %>% 
  group_by(Age, class, gender, survived) %>% count()
myNODE <- c(as.character(unique(myedge$Age)),
      as.character(unique(myedge$class)),
      as.character(unique(myedge$gender)),
      as.character(unique(myedge$survived)))
myNODE <- as.data.frame(myNODE) %>% 
  rowid_to_column('id') %>% mutate('id'=id-1)
myNODE$myNODE <- as.character(myNODE$myNODE)
myNODE <- as.data.frame(myNODE)

myedge2 <- titanic %>% 
  group_by(Age,class) %>% count() %>% 
  left_join(myNODE, by = c('Age' = 'myNODE')) %>% 
  ungroup() %>% 
  rename(from = id) %>% select(-Age) %>% 
  left_join(myNODE, by = c('class' = 'myNODE')) %>% 
  rename(to = id) %>% select(-class)

myedge3 <- titanic %>% 
  group_by(class,gender) %>% count() %>% 
  left_join(myNODE, by = c('class' = 'myNODE')) %>% 
  ungroup() %>% 
  rename(from = id) %>% select(-class) %>% 
  left_join(myNODE, by = c('gender' = 'myNODE')) %>% 
  rename(to = id) %>% select(-gender)

myedge4 <- titanic %>% 
  group_by(gender,survived) %>% count() %>% 
  left_join(myNODE, by = c('gender' = 'myNODE')) %>% 
  ungroup() %>% 
  rename(from = id) %>% select(-gender) %>% 
  left_join(myNODE, by = c('survived' = 'myNODE')) %>% 
  rename(to = id) %>% select(-survived)

myEDGE <- rbind(myedge2,myedge3,myedge4)

sankeyNetwork(Links = myEDGE, Nodes = myNODE,
              Source = "from", Target = "to",
              Value = "n", NodeID = "myNODE",
              fontSize = 10, nodeWidth = 10)
```




















