---
title: "quick example of latent profile analysis in r"
author: "jmpark"
date: "2019년 5월 2일"
output: 
  html_document :
    df_print : paged
---

[Quick example of Latent Profile Analysis in R](https://www.r-bloggers.com/quick-example-of-latent-profile-analysis-in-r/)  

* LPA(Latent Profile Analysis)
    + identify clusters of individuals based on responses to a series of continuous variables
    + LPA assumes that there are unobserved latent profiles that generate patterns of responses on indicator items  

* Data
    + [Young people survey](https://www.kaggle.com/miroslavsabo/young-people-survey)
    + Explore the preferences, interests, habits, opinions, and fears of young people
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
survey <- read_csv("https://raw.githubusercontent.com/whipson/tidytuesday/master/young_people.csv") %>%
  select(History:Pets)
dim(survey)
head(survey)
```

* There may be careless responding (eg. participants who selected the same value over and over)
    + use the `careless` package
```{r, message=FALSE, warning=FALSE}
library(careless)
library(psych)

interests <- survey %>%
  mutate(string = longstring(.)) %>%
  mutate(md = outlier(., plot = FALSE))
```

* `careless::longstring` : identifies the longest string of identical consecutive responses for each observation
* `psych::outlier` : find and graph Mahalanobis squared distances to detect outliers

```{r}
head(interests)
```

* cap string responding to a maximum of 10 
* use a Mahalanobis D cutoff of alpha = .001.
```{r, message=FALSE, warning=FALSE}
cutoff <- (qchisq(p = 1 - .001, df = ncol(interests)))
cutoff
interests_clean <- interests %>%
  filter(string <= 10,
         md < cutoff) %>%
  select(-string, -md)
dim(interests_clean)
head(interests_clean)
```

* `mclust`
     + performs various types of model-based clustering and dimension reduction
     + requires complete data (no missing)
     
```{r, message=FALSE, warning=FALSE, fig.align='center'}
library(mclust)

interests_clustering <- interests_clean %>%
  na.omit() %>%
  mutate_all(list(scale))  # scale to see the differences more clearly
head(interests_clustering)

BIC <- mclustBIC(interests_clustering)
plot(BIC)
```

```{r, message=FALSE, warning=FALSE}
mclustModelNames
summary(BIC)
mod1 <- Mclust(interests_clustering, modelNames = "VEE", G = 3, x = BIC)
summary(mod1)
```

* ICL(Integrated Completed Likelihood)
    + adds a penalty on solutions with greater *entropy* or classification uncertainty
```{r, message=FALSE, warning=FALSE}
ICL <- mclustICL(interests_clustering)
plot(ICL)
summary(ICL)
```

* BLRT(Bootstrap Likelihood Ratio Test)
    + compares model fit between *k-1* and *k* cluster models
```{r, message=FALSE, warning=FALSE}
mclustBootstrapLRT(interests_clustering, modelName = "VEE")
```

### Visualizing LPA
```{r, message=FALSE, warning=FALSE, fig.align='center', fig.width=9, fig.height=6}
library(reshape2)
means <- data.frame(mod1$parameters$mean, stringsAsFactors = FALSE) %>%
  rownames_to_column() %>%
  rename(Interest = rowname) %>%
  melt(id.vars = "Interest", variable.name = "Profile", value.name = "Mean") %>%
  mutate(Mean = round(Mean, 2),
         Mean = ifelse(Mean > 1, 1, Mean))
round(table(mod1$classification)/nrow(interests_clustering),2)
means

means %>%
  ggplot(aes(Interest, Mean, group = Profile, color = Profile)) +
  geom_point(size = 2.25) +
  geom_line(size = 1.25) +
  scale_x_discrete(limits = c("Active sport", "Adrenaline sports", "Passive sport",
                              "Countryside, outdoors", "Gardening", "Cars",
                              "Art exhibitions", "Dancing", "Musical instruments",
                              "Theatre", "Writing", "Reading","Geography", 
                              "History", "Law", "Politics", "Psychology", "Religion", 
                              "Foreign languages", "Biology", "Chemistry", "Mathematics", 
                              "Medicine", "Physics", "Science and technology",
                              "Internet", "PC","Celebrities", "Economy Management",
                              "Fun with friends", "Shopping", "Pets")) +
  labs(x = NULL, y = "Standardized mean interest") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top")
```

* red group : interested in science
* blue group : interested in arts and humanities
* green group : disinterested in both science and art, but moderately interested in other things

```{r, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
data <- data.frame(mod1$parameters$mean, stringsAsFactors = FALSE) %>%
  rownames_to_column() %>%
  rename(Interest = rowname) %>%
  mutate(X1 = round(X1, 2),
         X1 = ifelse(X1 > 1, 1, X1),
         X2 = round(X2, 2),
         X2 = ifelse(X2 > 1, 1, X2),
         X3 = round(X3, 2),
         X3 = ifelse(X3 > 1, 1, X3))
         
data %>%
  mutate(
    X1 = cell_spec(X1, "html", color = ifelse(X1 > 0.3, "red", "black")),
    X2 = cell_spec(X2, "html", color = ifelse(X2 > 0.3, "red", "black")),
    X3 = cell_spec(X3, "html", color = ifelse(X3 > 0.3, "red", "black"))) %>%
  kable(format = "html", escape = F) %>%
  kable_styling("striped", full_width = F)
```

```{r, message=FALSE, warning=FALSE, fig.align='center', fig.width=9, fig.height=6}
p <- means %>%
  mutate(Profile = recode(Profile, 
                          X1 = "Science: 16%",
                          X2 = "Disinterest: 60%",
                          X3 = "Arts & Humanities: 24%")) %>%
  ggplot(aes(Interest, Mean, group = Profile, color = Profile)) +
  geom_point(size = 2.25) +
  geom_line(size = 1.25) +
  scale_x_discrete(limits = c("Active sport", "Adrenaline sports", "Passive sport",
                              "Countryside, outdoors", "Gardening", "Cars",
                              "Art exhibitions", "Dancing", "Musical instruments", 
                              "Theatre", "Writing", "Reading", "Geography",
                              "History", "Law", "Politics", "Psychology", "Religion",
                              "Foreign languages", "Biology", "Chemistry", "Mathematics", 
                              "Medicine", "Physics", "Science and technology",
                              "Internet", "PC", "Celebrities", "Economy Management",
                              "Fun with friends", "Shopping", "Pets")) +
  labs(x = NULL, y = "Standardized mean interest") +
  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top")

p
```

* interactive plot
```{r, message=FALSE, warning=FALSE, fig.align='center', fig.width=9, fig.height=6}
library(plotly)
ggplotly(p, tooltip = c("Interest", "Mean")) %>%
  layout(legend = list(orientation = "h", y = 1.2))
```

























